---
title: "Towards caring 15-minute neighbourhoods"
author:
  - name: Anastasia Soukhov 
    email: soukhoa@mcmaster.ca
    affiliations:
        - name: McMaster University
          department: School of Earth, Environment & Society
          city: Hamilton, Canada
    attributes:
        corresponding: true
  - name: Léa Ravensbergen
    email: ravensbl@mcmaster.ca
    affiliations:
        - name: McMaster University
          department: School of Earth, Environment & Society
          city: Hamilton, Canada
  - name: Lucía Mejía Dorantes
    email: lucia.mejia.dorantes@gmail.com
    affiliations:
        - name: Consultant
          city: Karlsruhe, Germany
  - name: Antonio Páez
    email: paezha@mcmaster.ca
    affiliations:
        - name: McMaster University
          department: School of Earth, Environment & Society
          city: Hamilton, Canada
format: 
  docx
  # elsevier-pdf:
  #   keep-tex: true
  #   include-in-header:
  #     text: |
  #       \usepackage{graphicx}
  #       \usepackage{unicode-math}
  #       \usepackage{hyperref}
  #       \def\tightlist{}
  #       \usepackage{setspace}
  #       \doublespacing
  #       \usepackage{lineno}
  #       \linenumbers
  #   journal:
  #     name: 7th International Conference on Women and Gender in Transportation (WGiT)
  #     formatting: preprint
  #     model: 3p
  #     cite-style: authoryear
keywords: ["Accessibility", "Mobility of Care", "Gender", "15-minute neighbourhood", "Care Complete Neighbourhoods"]
bibliography: bibliography.bib
---

<!-- 

•	Establishing definitive research questions and the work’s contributions. Us deciding on this together will help me sculpt from what’s there
  •	Thoughts on methods? Can we improve presentation?
  •	Thoughts on results? Can we improve presentation? What are insights that are significant to you? What do you want to know more about?
o	Thoughts on how I weighted care destinations
o	Thoughts on aggregation by ‘care category’ vs ‘specific type’ of destination
o	Thoughts on additional dimensions – i.e., the inclusion of results that only consider ‘grocery, errand, and health centric’ destinations (for people who have no dependents). For the time being, I left this consideration out and just demo _all_ destinations. Also, what about LICO-AT discussion?
  •	Critical thoughts on how to tie walking and cycling modes together. For now, I did not include cycling results, though I have the cycling travel time matrices. We should use the cycling results perhaps to demo how the range of care access can be extended for certain neighbourhoods. A potential ‘what-if’ scenario?
  •	Next steps

-->

```{r}
#| label: setup-knitr-options
#| include: false

knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      cache=FALSE)

```
```{r }
#| label: set-up-copy-bibliography
#| include: false

# WARNING: DO NOT MODIFY THE BIBLIOGRAPHY ON THIS FOLDER, WORK WITH THE ONE IN FOLDER ./bibliography/
file.copy(glue::glue(here::here(), "/bibliography/bibliography.bib"), 
          getwd(),
          overwrite = TRUE)
```

```{r}
#| label: function for creating comments

word_comment <- function(comment, highlight = "") {
  if (isTRUE(knitr:::pandoc_to() == "docx")) {
    paste0('[', comment, ']{.comment-start id="0" author="Anastasia"',
           'date="2024-05-16T15:00:00Z"}', highlight, '[]{.comment-end id="0"}')
  }
}
```

```{r library-setup}
library(biscale)
library(cowplot)
library(magick)
library(dplyr)
library(flextable)
library(ftExtra)
library(glue)
library(here)
library(mapview)
library(pander)
library(scales)
library(sf)
library(tmap)
library(tmaptools)
```

```{r}
load(glue::glue(here::here(), "/data/care_dest.rda"))
```

```{r reformat care-dest specific categories}
care_dest <- care_dest |> 
  mutate(Care_Category_Specific = case_when(
    Care_Category_Specific  == "Daycare_EarlyON" ~ "Daycare or EarlyON",
    Care_Category_Specific == "Community_Centre" ~ "Community or Recreation Centre",
    Care_Category_Specific == "Daycare_Centre" ~ "Daycare or EarlyON",
    Care_Category_Specific == "Recreation_Centre" ~ "Community or Recreation Centre",
    Care_Category_Specific == "School" ~ "School",
    Care_Category_Specific == "Long-Term_Care_Home" ~ "Long-Term Care or Retirement Home",
    Care_Category_Specific == "Retirement_Home" ~ "Long-Term Care or Retirement Home",
    Care_Category_Specific == "Senior_Centre" ~ "Senior Centre",
    Care_Category_Specific == "Seniors_Active_Living_Centre" ~ "Senior Centre",
    Care_Category_Specific == "Post_Office" ~ "Post Office",
    Care_Category_Specific == "Library" ~ "Library",
    Care_Category_Specific == "Banks" ~ "Bank or ATM",
    Care_Category_Specific == "Grocer_ConvenienceStore" ~ "Convenience Store",
    Care_Category_Specific == "Clinics" ~ "Hospital or Clinic",
    Care_Category_Specific == "Grocer_Retail" ~ "Grocer",
    Care_Category_Specific == "Dentist" ~ "Dentist",
    Care_Category_Specific == "Hosptial" ~ "Hospital or Clinic",
    Care_Category_Specific == "Hospital" ~ "Hospital or Clinic",
    Care_Category_Specific == "Pharmacy" ~ "Pharmacy",
    Care_Category_Specific == "Park" ~ "Park"),
  Care_Category = case_when(
    Care_Category == "Child_Centric" ~ "Dependent-centric",
    Care_Category == "Elder_Centric" ~ "Dependent-centric",
    Care_Category == "Errand_Centric" ~ "Errand-centric",
    Care_Category == "Grocery_Centric" ~ "Grocery-centric",
    Care_Category == "Health_Centric" ~ "Health-centric"))

care_dest <- care_dest |> rename("care_ID" = "ID") |> mutate(care_ID = as.character(care_ID)) 
# care_dest|> st_drop_geometry() |> group_by(Care_Category) |> summarise(count = n())
```

 
# Abstract 

The 15-minute neighbourhood is a concept in planning that promotes urban spaces that are responsive to human needs and environmental sensibilities. These neighbuorhoods aim to accomplish this by ensuring that most daily necessities are not further away than a 15 minutes active trip [@Allam2022]. In practice, 15-minute neighborhoods are a form of normative accessibility [@paez_measuring_2012], and therefore related to accessibility analysis, an increasingly important tool that planners use to foster just and sustainable cities [@vale2023accessibility; @handyAccessibilityIdeaWhose2020]. Both the 15-minute neighborhood and accessibility measures are flexible enough to consider all destination types holistically, however, gendered examinations have been lacking in the literature. For instance, accessibility analyses have historically been employment-centric even when adopting a gendered perspective [@paez_jobs_2024]. 

To address this gender bias, this study investigates the 15-minute neighbourhood through the lens of Mobility of Care [@sanchezdemadariagaMobilityCareIntroducing2013]. This framework emphasizes the importance of travel related to unpaid work (care trips) in contrast to the better-studied travel to employment and leisure [@lawWomenTransportNew1999; @cresswell_gendered_2008; @Oestergaard1992; @levy1991towards; @levy2013travel]. While these trip types (work, care, and leisure) are essential, care trips are often relatively shorter-distance, proximate to residential/work/school, and comprise approximately 30% of adults’ daily trips [@sanchezdemadariagaMeasuringMobilitiesCare2019; @ravensbergen2023exploratory; @mejia2021mobility], thus fitting neatly within the concept of 15-minute neighbourhoods.

Our study casts the 15-minute neighbourhood onto the Mobility of Care framework. Empirically, we identify which areas in Hamilton, Canada, are 'caring 15-minute neighbourhoods'. To do so, a database of care destinations was compiled. In this database, care destinations include all places associated with sustaining household tasks needed for the reproduction of life namely: shopping (e.g., groceries), errands (e.g., libraries), health (e.g., dentist), and caring for dependents (e.g., schools). This database is used to estimate the number and mix of care-destinations that can be reached within a 15-minute walk-sheds from census centroids. We generate typologies to illustrate which neighbourhoods can and to what degree facilitate 15-minute access to care.

Our results indicate which 15-minute neighbourhoods are almost 'care-complete', that is, contain a sufficiently high mix of care destinations. Very few of these are found outside of the downtown core. Our study frames these neighbourhoods on a continuum of 'caring' and helps identify areas for policy intervention. Taken together, this study provides a theoretical bridge to connect and Mobility of Care, 15-minute neighbourhoods, and accessibility analysis. This should help to inform policy choices aimed at encouraging more just and sustainable mobility. 

# Introduction

Caring is one of the most fundamental activities in societies, since it consists of numerous activities that allow them to thrive. Caring activities are those that strive to meet the physical, psychological, and emotional needs of others [@ilo_care_2018]. Yet, it is one of the most unequal, undervalued, and even devalued activities worldwide. Traditionally, caring activities have been borne on women’s shoulders [@hayden_grand_2000; @hochschild_second_2012]. According to @ilo_care_2018, women and girls perform more than three-quarters of the total amount of unpaid care, a gender gap that varies geographically [@RN5]. 

This unequal share of caring responsibilities leads to multifaceted gendered differences: in career development, profession selection, contract type, pay gap, and time poverty, as recognized by various international organisations [@EIGE2016; @ilo_care_2018]. In fact, early feminists in the United States had already observed that domestic labor was the single most relevant cause of women’s inequality in society [@gordon_womans_1976; @hayden_grand_2000].

Care also has important implications for spatial and transportation planning. Indeed, researchers from various fields have highlighted that western modern industrial societies have imposed a physical separation of the private sphere from the public one, splitting production and reproduction, public and private spaces or simply work and home, concentrating on the economic outcome of being mobile [@cresswell_gendered_2008; @lawWomenTransportNew1999; @levy2013travel; @little_gender_1994; @tronto_toward_1990]. The man with legal rights, paid labour public sphere and woman in the private sphere, dependent, with family duties, personal relations and unpaid labour [@tronto_toward_1990]. In this line, @cresswell_gendered_2008 described how masculinity was related to mobile and active, whereas femininity to static. As an example of this dichotomy, transport planning historically framed non-work/school trips as 'non-mandatory' activities, downgrading these activities in comparison to work. In more recent years, planners have begun to resist this spatial separation of the public and private sphere to promote for sustainable cities. Proximity planning has become more normalized, at popular example of which is the 15-minute neighbourhood, in which residents can reach all necessary destinations within a 15-minute walk. The 15-minute neighbourhood concept, however, has yet to foreground care.

While much care labour takes place in the private and feminised sphere, completing care activities requires one to be mobile. In other words: caring for oneself and others requires trips to, amongst other places, grocery stores, day cares, elder care, and health care destinations. 

From this research motivation, Sánchez de Madariaga coined the term ‘mobility of care’ in 2013 to refer to this travel or all travel required to sustain the needs of a household, such as grocery shopping, escorting children, travelling to health appointments, and running errands [@sanchezdemadariagaMobilityCareIntroducing2013]. While decades of research had examined these unique household-serving trips, her seminal work was the first to consider all these trips as one category. In doing so, she showed how mobility of care is a significant proportion of daily travel. Indeed, in her original work, and in the studies that have followed, it appears that approximately 30% of adults’ trips are for care purposes [@sanchezdemadariagaMobilityCareIntroducing2013; @sanchezdemadariagaMeasuringMobilitiesCare2019; @ravensbergen2023exploratory]. Perhaps unsurprisingly given the gendered division of care work discussed above, women have consistently been found to complete more mobility of care trips than men. In one study, mobility of care comprised 32% of women’s daily trips compared to 28% of men’s. While this gap was significant, it was found to be far greater in lower income households where women complete 20% more care trips than men [@ravensbergen2023exploratory]. Sánchez de Madariaga not only shows how important these mobility of care trips are, but also highlights the ways in which “mobility of care is systematically under-represented in any analysis of urban transport” (p. 37). Transport surveys and tools do not directly capture mobility of care, which re-enforces the idea that these trips are not a significant part of daily mobility. 

Cities have simplified individuals as a universal and neutral entity. As a result, public spaces are differently experienced depending on our multiple identities. Indeed, asymmetric in the distribution of time of citizens is deeply anchored with the public and private spaces organization within cities. In this light, we bring together the literatures on 15-minute cities and mobility of care by focusing on the potential access to destinations of care through an empirical case study of the mid-sized city of Hamilton, Canada. We acknowledge that caring is composed by actors and activities but aim to spotlight the role that the urban environment plays on our daily lives, on easing our participation and on exercising our rights to the city as individuals with multiple identities. Our work hopes to contribute to a fruitful discussion that may trigger urban and transport policy interventions that increase gender-equity in caring activities among all actors.

## Feminist 15-minute neighbourhoods

`r word_comment("TO ADD: some good feminist 15 minute cities. It's novel but movements exists in the Canadian context i.e., https://x.com/brenttoderian/status/1800371515877384527?s=48&t=QI0WSvV_FDb82ZGeLdJ9lQ. Also discuss Barcelona and Bogota (care blocks)", " ")`

Within the backdrop of urban planning and 'mobility of care', we connect our work to the 15-minute neighbourhood concept. 

The spatial structure of cities and regions has been described in different ways and its advantages and disadvantages has been discussed with regard to a wide variety of issues, for example, regarding health, the environment, economic, social and equity impacts. After the traditional monocentric vision of the city with a core Centre Business District (CBD), other alternatives appeared, especially due to the expansion of cities caused by the automobile and motorized transport. The polycentric urban planning emerged based on the idea of sub-centers (or multiple business centers) with their own services and functions, well-connected among other surrounding areas, where the links between land-use and transport were optimised. Economic activities (such as retail or commercial office development) and job centers played an important role in shaping the spatial structure [@schneider_transit_1981; @sweet_are_2017]. This concept may have different interpretations depending on the spatial and geographical scales [@hall_polycentricity_2009].

In the last decades, the need for more sustainable, healthier, and liveable cities has received more attention. Planners and policy makers have proposed different alternatives where residents reach “essential activities” within walkable or cycling distances, reducing the need for using motorised transportation modes. The 15-minutes cities or "x"-minutes cities, have been long under the spotlight. This approach has also received the name of Chrono-urbanism or the “cities of proximity” [@logan_xminute_2022; @moreno_introducing_2021]. 

In the international context, there are different examples that highlight the importance of good access to different services and amenities within walkable distances. Probably one of the first cities to adopt the proximity approach was Portland. The Portland Plan [@city_of_portland_20minute_2010] was adopted by the City Council on April 25, 2012 aiming at fostering an inclusive city development based on prosperity, education, health, and equity for the following 30 years. The progress report states that equity should be achieved, an Office of Equity and Human Rights was established focusing on equitable service delivery to all residents. Nonetheless the progress report only measures progress on topics related to racial equity or people with disabilities [@portland_government_portland_2017]. `r word_comment("NOTE TO SELF: And so what? Is this a research gap?", " ")`

In Paris, the adoption of the 15-minute neighbourhood put into the spotlight this approach. They targeted six social functions that had to be easily accessible at any part of the city: housing, work, health care, groceries, education and leisure [@ville_de_paris_paris_2022]. In 2021, Ottawa also adopted a 15-Minutes approach in the new City of Ottawa Official Plan [@ville_dottawa_quartier_2021]. More recently, @moreno_introducing_2021 considers four dimensions on how the 15-minute neighbourhood framework: density (in terms of people per sq.km); diversity (including mixed-land use and diversity of people; the temporal and spatial proximity to essential services (residential areas, work, commercial areas, education centres, health facilities and other basic institutions) and digitalisation.  In this respect, the feminist perspective of the cities of proximity is still underestimated. Only few examples exist that specifically target this issue [@gil_sola_choose_2022; @macintyre_her_2022]. Given the impact of caring activities on everyday life, the chrono-urbanism should specifically target the analysis of caring activities.

## Mobility of care and mode-constraints 

Given that almost one third of daily trips are for care purposes [@sanchezdemadariagaMobilityCareIntroducing2013], it is important to plan for mobility of care. There are also equity considerations to consider; women, and especially low-income women, are particularly in charge of these duties and therefore, complete this travel more than men. Planning for mobility of care may also warrant different considerations because mobility of care trip characteristics differ from other trips, notably those of the commute to work. For instance, many of these trips are made through trip-chains, i.e., when a non-work-related stop is made during the commute. Women have been found to do so more than men, which is understood as being due to their need to balance unpaid-household serving care and paid employment (e.g., picking up groceries or children on the way home from work) [@ravensbergen2023exploratory; @scheinerWomenComplexDaily2015].

Further, compared to the trip to work, mobility of care trips are more frequently completed by foot, and less frequently done by transit or bicycle [@ravensbergen2023exploratory]. Further, evidence suggests that walking is more common for care than for work trips because care trips tend to be short [@ravensbergen2023exploratory]; if care destinations are closer to the home, it is more likely that these destinations are walkable. Ample literature examines the benefits and barriers to walking care destinations, such as schools (e.g., [@RN7;@RN8;@RN9]) and grocery stores (e.g., [@RN10;@RN11]). However, little of this work uses the mobility of care framework. Instead, they tend to consider singular care destinations in research focusing on walkability. 


`r word_comment("TO ADD: Replace the following three paragrapgs with a more elaborated discussion on walking as the primary mode for 15-min cities and how it overlaps with mobility of care. ", ".")`
In car dependent cities, cycling commute rates tend to be low. Emerging evidence has found that they are even lower for care trips than for work trips. For instance, in Montréal it is estimated that 2% of work trips completed by adults are done by bicycle, while only 1.1% of care trips are done using this sustainable travel mode [@ravensbergen2023exploratory]. Qualitative research has also found that in car-dependent contexts, many consider it inappropriate and unsafe to conduct escort trips by bicycle [@ravensbergenVelomobilitiesCareLowcycling2020; @sersliRidingAloneTogether2020]. Then again, other studies have uncovered the joys inherent in travelling by bicycle with children [@RN14]. This qualitative evidence has also found that the bicycle is excellent for trip-chaining owning to its flexibility and ease of parking, and that grocery shopping can be done by bicycle, though this can require some additional materials (e.g., paniers) and competencies (e.g., knowing where to put your groceries)[@ravensbergenVelomobilitiesCareLowcycling2020]. Research in the Netherlands, a high-cycling context, has found that parents can and do use their bicycles to complete mobilities of care. Indeed, one study compared women with children’s bicycle behaviours to those without children and found that the having a child did not change the amount of cycling done, but the characteristics of cycling trips: mothers’ trips tended to center around their childrens’ needs [@RN15]. Taken together, it is evident that the bicycle has potential to serve mobility of care: it is flexible, low-cost, and used for these trips in many contexts. Other barriers to this travel mode seems to exist in low-cycling contexts. 

The difference in uptake of public transport for work compared to care trips is staggering. In Montréal, it is estimated that 18.7% of trips to work are made by transit. In contrast only 5.8% of care trips are made using transit [@ravensbergen2023exploratory]. While research examining accessibility to care destinations by transit exists (e.g., [@RN16;@RN17;@RN18;@RN19]), little research examines why few people use this sustainable transport mode for mobility of care. Though transit use for mobility of care is low, one study found that is it is significantly higher for low-income women. Indeed, approximately 16% of low-income women in Montreal were found to use transit for care trips (compared to 4% of high income women) [@ravensbergen2023exploratory]. In many cities, public transport is not designed to meet household demands in a convenient way [@RN20;@grantsmithManagingChallengesCombining2016]. Given how low transit use is for care trips amongst the general population, it is possible that many of these low-women using transit are transit-dependent. In other words, they may be using transit for care trips because they do not have access to a car, and not because transit is their preferred travel mode. 

It appears that the car is the preferred travel mode for mobility of care. In one study, car use was higher for mobility of care trips than the commute to work (82.3% vs. 74.7%) [@ravensbergen2023exploratory]. Further, the rate of car use is higher amongst trips that are part of a chain than those that are not (85.28% vs. 78.09%) [@ravensbergen2023exploratory]. Some studies found that a car is viewed as the easier way to incorporate unpaid mobility of care into one’s daily travel as it provides convenience and mitigates concerns about traffic and stranger danger [@maciejewskaHaveChildrenThus2019, @carverParentalChauffeursWhat2013]. Others, however, have found that women are more likely to complete household-serving travel duties by alternative modes than men due to their lower car access [@RN24;@RN25]. It seems that car use may be preferred for mobility of care trips, but planning mobility of care for the car may leave those who do not have access to a car, many of them low-income women, with difficult mobility of care trips. 

The discussion around mobility of care and travel mode may also be relevant to larger spatial trends, including gendered spatial entrapment. Feminist geographers have noted for decades the spatial impact of women’s disproportionate household-serving labour. A relationship has been noted whereby many women’s household responsibilities results in job searches that emphasize job opportunities that are near the home and part-time or with flexible hours to allow for care responsibilities. This leads to a lower probability of securing higher paid, and often further afield, employment [@hansonGenderMobilityNew2010;@lawWomenTransportNew1999;@RN28;@RN29]. This trend has been found to be surprisingly resilient, with improvements only taking places slowly [@RN30]. Evidence has also shown that this spatial entrapment can be more in car dependent contexts, as access to cars impacts employment rates [@RN31], and is lower amongst women, and especially women of colour [@RN32].

Taken together, there appears to be a preference for travel by car for mobility of care trips. This is problematic, however, in at least two major ways. First, cars are less sustainable than other modes of transport. Normalizing car use for care trips does not advance the low-carbon transport transition that most cities are aiming for (at least in principle!). Since care trips tend to be short, they are also perfect candidate to be completed by bus, bicycle, or by foot. That is, if these alternative and sustainable modes are planned with care in mind. Second, there is equity concerns with assuming people are accessing care destinations by car. Many people, and low-income women in particular, do not have access to a car. Some are also experiencing a degree of spatial entrapment. Having safe and convenient access to care destinations using alternatives to the private car is especially important for those who must rely on these options regardless of whether they are well planned. 

The objectives of the paper are as follows: 

1)	City planning is moving to a people-centered approach focusing on accessibility to basic needs. However, chrono-urbanism stems from conventional methods that overlooks mobility of care perspectives. This paper presents a 15-minute cities methodological approach to assess mobility of care.
2)	We develop this approach using a mid-size typical Canadian city and a novel care destination dataset as case study
3)	Furthermore, the potential accessibility to and diversity of care destinations in the case study is analysed. We identify to what degree are neighbourhoods caring and/or complete. To do so, we make use of self organizing maps (SOM), an unsupervised machine learning data-dimensionality reduction technique to identify caring and complete neighbourhoods.
4)	We then apply the same SOM methodology to determine profiles of population groups to understand who enjoys these neighbourhoods and who does not.

# Data and Methods

## Case study context

In Canada, women have become more educated and have an achieved an increased participation in the labour market [@government_of_canada_gender_2023]. In fact, the educational level of women has surpassed most men [@government_of_canada_women_2023]. Reports observe that the gender pay gap has narrowed but disparities persist [@drolet_intersectional_2023;@pay_equity_office_gender_2023]. According to the statistics, in Ontario between 1998 to 2022 the gender gap narrowed but still women make 87 cents on average for every dollar made by a man [@pay_equity_office_gender_2023]. All reports observe that it affects women differently, women with disabilities, immigrants, indigenous or racialized suffer especially from the gender pay gap. Similarly to other countries worldwide, the occupational segregation remains and women are segregated into lower paid job classes and there is a dominance of women in minimum wage jobs [@drolet_intersectional_2023]. Furthermore, women are underemployed, and make up most part-time work. In this respect, the lack of childcare alternatives enforces this obligatory part-time work [@equal_pay_coalition_myths_2022]. All researchers and organisations agree that an intersectional approach is of utmost importance to better plan policies to enforce gender equality. 

This study focuses on Hamilton, Ontario, a mid-sized city on the shore of Lake Ontario. Hamilton has a heterogenous land-use, with a populated and dense urban core, surrounded by suburban development, which is itself surrounded by rural communities made possible through a greenbelt. In this work, we analyse the residential parcel centroids, approximately ~140,000 locations. We aggregated the points at the level of Canadian Census Dissemination Area (DA) along with the population and population per parcel plots in @fig-Fig1.

```{r plot-figure1, out.width=600, out.height=600}
#| label: fig-Fig1
#| fig-cap: "The number of residential parcels per DA in 2020 (top), the population (middle) retrieved from the 2021 Canadian Census, and the rate of population per parcel per DA (bottom). All scales in quartiles. Basemap shapefiles are sourced from the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics(glue::glue(here::here(), "/figures/pop_descriptive_plots_1.png"))
```

Hamilton also exhibits spatial disparities in social and economic indicators @fig-Fig2. The densely populated inner city is characterised by lower average incomes, and a higher prevalence of households living under the LICO. The suburban parts of the city tend to have a greater proportion of children and a lower proportion of one parent households.  

```{r plot-figure2, out.width=600, out.height=600}
#| label: fig-Fig2
#| fig-cap: "Socio-economic and demographic variables that characterise accessibility to care destinations retrieved from the 2021 Canadian Census. All scales in quartiles. Basemap shapefiles are sourced from the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics(glue::glue(here::here(), "/figures/pop_descriptive_plots_2.png"))
```

## Estimated travel times

Here we will describe the estimated 15-minute walk travel times using {r5r}, from all residential parcel points to all care destinations assuming default walking speed (3.6km/hr) and OSM road network for Hamilton. `r word_comment("TO ADD: add summary statistics", " ")`

## Care destination dataset

A spatial dataset of care destinations for Hamilton was compiled; sources include provincial and municipal open data portals [@governmentofontarioOntarioDataCatalogue2023; @cityofhamiltonOpenData2023], Data Axle, a consumer dataset compiled of businesses and companies within Canada [@axledataConsumerData2023] and manual retrieved through Google Maps. Each destination was categorized based on the specific type of care being accessed following the mobility of care research by @sanchezdemadariagaMeasuringMobilitiesCare2019. Initial categories include child, elder,  grocery, health, and errand -centric destinations. For the purpose of this study, child- and elder- centric are combined to represent a 'dependent-centric' category. Category sources of data and preparation notes are detailed in @tbl-Tbl1. Their spatial distribution and sub-categories are visualised in @fig-Fig3.

```{r creating-table-content}
dat <- data.frame(matrix(c("Child-centric",
                           "[@opendatahamiltonEducationalInstitutions2022; @governmentofontarioOntarioDataCatalogue2023; @cityofhamiltonChildCareRegistry2023; @opendatahamiltonParksDatabase2022; @opendatahamiltonRecreationCommunityCentres2022]",
                           "Schools, daycares, and community centres, recreation centres, and parks: 1,190 locations are included. After manual review, all locations that typically do not serve children were removed including: Post-Secondary, Adult-Learning Centres, Group Homes, and Foster Care Centres. Further, through examination some Section 23 institutions defined as _“centres for children who cannot attend school to meet the needs of care or treatment, and rehabilitation”_ [@governmentofontarioministryofeducationFundingEducationCommunity2023], were kept due to their innate connection to care.",
                           "Elder-centric",
                           "[@opendatahamiltonRecreationCommunityCentres2022; @ontarioministryofhealthgeohubMinistryHealthService2023]",
                           "Senior centres, long-term care homes, and retirement homes: 75 destinations are identified.",
                           "Grocery-centric",
                           "[@axledataConsumerData2023]",
                           "Grocery stores, namely a place a household could buy groceries ranging from convenience stores to large retail stores: 381 destinations are identified. Data is filtered by Company Name, Suite Number, Address, City, Province, Phone Number and Postal Code. The type was then identified e.g., grocers specialty foods, grocers retail, grocer health food, grocer wholesale, grocer curbside, grocer delicatessen wholesale, grocer convenience. Data was cross-referenced to ensure all included locations were operational and legitimate grocery stores.",
                           "Health-centric",
                           "[@ontarioministryofhealthgeohubMinistryHealthService2023; @hamiltonniagarahaldimandbranthealthlineWalkInMedicalClinics2023]",
                           "Hospitals, pharmacies, clinics, and dentist offices: 421 destinations are identified. Hospitals and pharmacies were retrieved while clinics and dentistry clinics were manually scraped from a healthcare services database and checked via Google Maps to remove non-operational locations and confirm dentistry-orientation. ",
                           "Errand-centric",
                           "Hamilton libraries [@opendatahamiltonLibraries2022], post office locations [@axledataConsumerData2023;@canadapostPostOfficeLocator2023], and datasets of all national bank chains [@bankofmontrealBankMontrealBranch2023; @thehongkongandshanghaibankingcorporationlimitedhsbcHSBCBranchABM2023; @nationalbankNationalBankBranches2023; @royalbankofcanadaRBCBranchABM2023; @scotiabankScotiabankABMBranch2023; @thetorontodominionbankTDBankBranch2023].",
                                                      "Libraries, post offices, and banks: 158 destinations are identified. Post offices are retrieved from a mix of databases, and duplicates are removed. Banks are also derived from Data Axle and then cross-referenced to ensure data quality with a Bank Locator website for all national banking firms."
                           ), ncol = 3, byrow = T))
```

```{r creating-table}
#| tbl-cap: "Details on the preparation and data sources of care destinations."
#| label: tbl-Tbl1
set_flextable_defaults(font.size= 10,
                       font.family = NULL,
                       text.align = 'justify',
                       table.layout = 'fixed')

flextable(dat) %>%
  ftExtra::colformat_md() %>%
  compose(
    i = 1, j = c(1,2,3), part = "header", value = as_paragraph(c("Care category", "Sources", "Data preparation notes"))) %>%
  width(j = 1, width = 1.7, unit = "cm") %>%
  width(j = 2, width = 4.3, unit = "cm") %>%
  width(j = 3, width = 11, unit = "cm")
```

```{r plot-figure3, out.width=600, out.height=600}
#| label: fig-Fig3
#| fig-cap: "The locations of care destinations in Hamilton separated by the author-generated categories of: dependent-, errand-, grocery- and health- centric care categories. Basemap shapefiles are sourced from the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics(glue::glue(here::here(), "/figures/care_categories_plot.png"))
```

## Methods
A 15-minute normative walk catchment is used to represent the 15-minute neighbourhood concept. This has been done in previous literature such as @vale2023accessibility, @graellsCityCitiesMeasuring2021 and @gaglioneUrbanAccessibility15minute2022. In this work, the cumulative opportunity score is calculated for each parcel $i$ and aggregated by care category $c$ and takes the following mathematical form:
$$
S^c_i=\sum_{j}O_j^{c}\cdot f(c_{ij})
$$
\noindent Where:

- $i$ is a set of parcel point origin locations.
- $j$ is a set of destination locations.
- $c$ is a set of care categories.
- $O^c_j$ is a number of opportunities of category $c$ at $j$.
- $c_{ij}$ is the travel cost between $i$ and $j$.
- $f(\cdot)$ is an impedance function of $c_{ij}$; within the cumulative opportunity approach, it is a binary function that takes the value of 1 if $c_{ij}$ is less than a selected value [@handyMeasuringAccessibilityExploration1997].
- $S_{i}^c$ is the unconstrained accessibility, the sum of weighted opportunities reachable within $f(\cdot)$, at each $i$ for each $c$.

*Entropy* measure has been widely used to characterise the diversity of land-use mix in early work by @frankLinkingObjectivelyMeasured2005 based on the species diversity Shannon-Wiener index that expresses relative evenness throughout a sequence (this measure among other species diversity indices of the time are reviewed in @whittaker1972evolution). It has also been used in social sciences to comprehend the occurrence of social events in a structured manner specially with regard to life trajectories [@ritschardSequenceAnalysisRelated2018]. 

In the realm of urban studies and planning literature the malleable concept of entropy has been used to characterise the diversity in land-use mix [@ewingTravelBuiltEnvironment2010a], especially in the context of active modes [@luUrbanDensityDiversity2017; @mavoaIdentifyingAppropriateLanduse2018] and suburban sprawl [@randallGISBasedLand2015]. As well, it has been used to understand mobility behaviour [@mcbride2020exploration; @montero2023applying; @montero2023role]. Work that has used diversity indices alongside accessibility analysis are scarce, though examples of its use as parameters within accessibility scores in the case of employment opportunities [@chengMeasuringUrbanJob2013; @daiIncorporatingJobDiversity2018] and more recently to describe the diversity in transit facilities [@yinIncorporatingFacilityDiversity2024] are present. To the authors knowledge, the use of an entropy measure to reflect the diversity in care destination amenity types and in context within a 15-minute neighbourhood, has yet to be done. 

As such, the entropy measure is used to reflect the diversity of care destinations categories. The value ranges from 0 to 1, with 1 representing total evenness in the number of care opportunities in each category that can be reached. $i$ is a parcel, $c$ is the category of care destinations, $n_c$ is the count of care destination categories (in this work 4), $S_i^c$ is the spatial accessibility of care destinations within a 15-minute walk for a parcel, and $D_i$ is the diversity. As $D_i$ represents evenness in categories, a parcel $i$ can be assigned a diversity of 1 if it has $S_i^c$=1 of each category or if it has $S_i^c$=100 of each category. Contrariwise, it can be assigned a low score if it has low but different scores for each category as well as if it has high but different scores for each category. 

$$
D_{i} = \frac{-\sum_{c} S_i^{c} / \sum_{c}S^{c}_{i} \times ln(S_i^{c} / \sum_{c}S^{c}_{i} )}{ln(n_c)}
$$

Self organizing map (**SOM**) is used to cluster the types of neighbourhoods by their magnitude in 15-minute walking accessibility to caring opportunities and diversity in these opportunities and profile these clusters based on their socio-economic characteristics. SOM is a type of artificial neural network (ANN) trained using unsupervised learning that produces a two-dimensional representation of the input space. It is a dimensionality reduction method as it applies competitive learning by way of a neighbourhood function thereby preserving the topological properties of the input space and producing combinations of the input space that we intuitively profile. It has been used in transportation related research, e.g., @victorianoTimeSpaceMoney2020. 

The SOM procedure is broken into 2 steps: first, five superclusters of the four $S_i^c$ scores and the diversity score $D_i$ are identified. The number of superclusters strike a balance between variance explained and supercluster interpretation in what we define as ranging from "not caring" to "very caring" and "not complete" to "complete". Second, the superclusters are profiled using a variety of socio-economic and demographic indicators that the mobility of care literature finds significant (@fig-Fig2).

# Results

Accessibility per category @fig-Fig4. The concentration of opportunities for all category types is concentrated in Hamilton Central - with the highest concentration being in the lower city and pockets of the high concentration on the mountain `r word_comment("(TO ADD: The escarpment line)", " ")`. Grocery-centric opportunities are most concentrated, followed by dependent-centric and health-centric. Errand-centric is the most sprawled. 

```{r plot-figure4, out.width=600, out.height=600}
#| label: fig-Fig4
#| fig-cap: "The number of care destinations that can be reached within 15 min walk per care category for a median parcel in each DA. Scales are in quartiles. These are four of five inputs into the superclusters. Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics(glue::glue(here::here(), "/figures/percats_copp_plot.png"))
```
Diversity @fig-Fig5. Concentrated in the downtown core.
```{r plot-figure5, out.width=600, out.height=600}
#| label: fig-Fig5
#| fig-cap: "The diversity measure (div) based on the proportion of care category spatial accessibility (figure 4). Scale is in tertiles. The fifth input into the superclusters. Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics(glue::glue(here::here(), "/figures/all_cats_entropy_copp_plot.png")) 
```

Superclusters based on accessibility per category and diversity (the two above figures). 5 clusters are identified, see their dendrogram (right) and grid map (left) in @fig-Fig6 `r word_comment("change colours of SC to match the following plots", ":")`.

```{r plot-figure6-create, eval=FALSE}
plot1 <- ggdraw() + draw_image(glue::glue(here::here(), "/figures/access_sc_plot_2.png"))
plot2 <- ggdraw() + draw_image(glue::glue(here::here(), "/figures/access_sc_plot_1.png"))
plot3 <- plot_grid(plot1, plot2)

ggsave(glue::glue(here::here(), "/figures/access_sc_plot_combo.png"), dpi=300)
```

```{r plot-figure6, out.width=800, out.height=800}
#| label: fig-Fig6
#| fig-cap: "The 5 resulting superclusters from the SOM output (left) and the associated dendrogram structure (right)."
#| 
knitr::include_graphics(glue::glue(here::here(), "/figures/access_sc_plot_combo.png")) 
```

The constitution of the superclusters in shown in the decision tree diagram in @fig-Fig7. Decision trees area a supervised learning technique, which can use the results of the clustering algorithm as labels, and the input variables as features. The clustering algorithm performed well as class membership is relatively homogeneous as the majority of bar plots consist of a supermajority of one type of supercluster class membership, i.e., supercluster 5 are parcels that have an Errand-centric access of less than 0.5 opportunities (bottom quartile), a diversity score of less than 0.277 (bottom tertile), and any Dependent-centric access score of greater than 0.5 (and can take on any value of the remaining input variables). This technique provides a data-driven approach to supercluster formation which we then attribute meaning to.

```{r plot-figure7, out.width=600, out.height=300}
#| label: fig-Fig7
#| fig-cap: "The decision tree profiles of the SOM output."
knitr::include_graphics(glue::glue(here::here(), "/figures/decision_tree_profiles_sc.png"))
```


The dendrogram demonstrating the supercluster class membership of each parcel in @fig-Fig7 is summarized, with our interpretation, as follows:
- Very caring (mean category accessibility between 4 to 18.6 and total accessibility median of 46) - very complete (diversity index of mean 0.9 and median 0.90) (SCluster 2)
- Caring (mean category accessibility between 2.1 to 9 and total accessibility median of 18) -  complete (div index of mean 0.81 and median 0.82) (SCluster 1)
- Somewhat caring (mean category accessibility between 0.2 to 6 and total accessibility median of 13) - complete div index of mean 0.71 and median 0.74) (SCluster 4)
- Somewhat caring (mean category accessibility between 0.2 to 6 and total accessibility median 7) - not complete (div index of mean 0.36 and median 0.40) (SCluster 3)
- Not caring (mean category accessibility between 0 to 4.7 and total accessibility median of 3) - not complete (div index of mean 0.02 and median 0) (SCluster 5)

To further visualize supercluster membership, we spatially demonstrate which supercluster best describes each parcel in @fig-Fig8 (top plot): the median parcel's supercluster membership per DA is tabulated and the majority membership is represented. Visually, the super clusters closely follow the quartile split of the accessibility to care destinations (all categories combined) and tertile split of the diversity scores (visualised in @fig-Fig5) `r word_comment("AN IDEA DISCUSSED: Overall, in @fig-Fig8 there are many pockets of good/great care accessibility+diversity of opportunities Hamilton. These are objectively caring neighbourhoods. However, people may subjectively disagree... ", ".")`. 

```{r plot-figure8, out.width=600, out.height=600}
#| label: fig-Fig8
#| fig-cap: "The median parcel supercluster membership (top). The parcel median total accessibility (bottom). The parcel m Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics(glue::glue(here::here(), "/figures/SCLab_and_total_copp_plot.png"))
```
To enhance the meaning of the superclusters beyond a spatial conceptualisation of "caring" and "complete", was _who?_ resides there. We implement another decision tree in which the population-weighted socio-demographic characteristics of the parcels (retrieved from the 2021 Canadian Census) are input variables and the resulting output are 'socio-domographic superclusters' of the caring/complete superclusters. The input variables included: % households below median household after-tax income, LICO prevalence, average number of children per household, % not in the labour force, Gini index on adjusted household after-tax income, % visible minority, % population aged 0 to 14, % single parent household, % who walk to work (relative to bike, care/truck/van, public transit and 'other'), % of owner household in core housing need (i.e., inadequate housing structure or paying higher than 30% income on housing), % of owner households with a mortgage, % of tenant households in subsidized housing, % of tenant households in core housing need, % no certification, diploma or degree (relative to population with high school, college, or university certificate, diploma or degree).

Of all the variables, those that proved to be significant in profiling the five superclusters were only four variables: % households below median household after-tax income (bottom income), % population aged 0 to 14, % single parent household, and LICO. See @fig-Fig9 that provides the splits. Unfortunately, the supercluster membership is not as homogeneous as the caring/complete supercluster dendrogram (@fig-Fig7). However, we can interpret the socio-demographic superclusters (five that result) as labels to tell a narrative of who may reside in what caring/complete supercluster. 

```{r plot-figure9, out.width=600, out.height=300}
#| label: fig-Fig9
#| fig-cap: "The dendrogram of the supercluster demographic profiles.."
knitr::include_graphics(glue::glue(here::here(), "/figures/access_sc_profiles.png"))
```
(NOTE: for reference, see the old bivariate map in the appendix for comparison)

In @fig-Fig9, discussing from right-to-left, parcels in DAs with 45% of the population below the median or lower (richer) are often classified with poor accessibility: either "Somewhat caring and not complete" SC3 or "Not caring and not complete" SC5, where parcels within DAs with a higher proportion of single parent households (13% and above) are most often SC3 and below are most often SC5. For parcels with higher than 45% of bottom-income (poorer), those with a high proportion of children (age 0 to 14) are also likely to be within the "Somewhat caring and not complete" SC3 - just like parcels in richer and more-single-parent DA counterparts. 

These findings related to potential child-care access are notable: single parents provide more care and hence may be more likely to choose, if they are able, more accessible neighbourhoods than the rest of the population in the same household income grouping. In a Canadian and U.S. context, households with children may locate to catchments with 'better' schools, often more suburban with lower potential access scores overall. Furthermore, statistics show that in the last years, the number of single parents families in Canada has increased. In 2022, there were about 1.84 million single parents compared to about 1.56 million in 2010 and 1,3 million in 2001 (https://www.statista.com/statistics/443342/single-parent-families-in-canada/ & https://www.justice.gc.ca/eng/rp-pr/jr/jr12/p5a.html#ftn4). Furthermore, Ontario is the Canadian province where most of the lone parent families live (749,769 families, https://www.statista.com/statistics/446667/lone-parent-families-in-canada-by-province/). Most lone-families are headed by female-lone parent. (AS TO ADD IN HAMILTON SPECIFIC NUMBERS)

Interestingly, parcels in DAs with lower than 13% children population demonstrate a better accessibility to care destination story. Parcels within DAs that constitute more than 45% population below median household income (poorer) - especially above 9.8% LICO (low income cut-off prevalence - likely to be the poorest) are often within "Very caring and very complete" SC1. Parcels located in DAs with the same lower median income and child population but have _lower_ than 9.8% LICO (not the poorest) also have good accessibility, as they are most often in "Caring and complete" SC1 as well as being in most often in the "Somewhat caring and complete" SC4 relative to other parcels.

Our findings illustrate differences in patterns of 15-minute walk accessibility to care destinations most prominently profiled by income, low-income prevalence, children and single-parent status.

 `r word_comment("IDEAS DISCUSSED: regarding @fig-Fig9 - 1) We know that highest accessibility & complete is downtown. But people there have fewer children + lowest income are downtown.  On the horizon as rent increases,  the most precarious may be pushed out and replaced by people who can pay higher rents. 2) The lowest accessibility & most incomplete are suburban -- they are likely higher income. Interesting to note: one-parent families are more likely to be time poor -- and leaving in low access/incomplete suburbs puts more at risk. We know a think or 2 about single parents and jobs -- access to jobs was high where single parents were not. (Catherine, farber Paez paper). 3) Pockets of the suburbs have high access to care. The suburbs do something -- they have care amenities. ", " ")`
  `r word_comment("IDEAS DISCUSSED: regarding @fig-Fig9 - 2) The lowest accessibility & most incomplete are suburban -- they are likely higher income. Interesting to note: one-parent families are more likely to be time poor -- and leaving in low access/incomplete suburbs puts more at risk. We know a think or 2 about single parents and jobs -- access to jobs was high where single parents were not. (Catherine, farber Paez paper). ", ":")`. 

 `r word_comment("IDEAS DISCUSSED: regarding @fig-Fig9 - 3) Pockets of the suburbs have high access to care. The suburbs do something -- they have care amenities. ", ":")`. 


# Conclusions and Discussion

There seems to be a spatial mismatch where dwelling availability and affordability play an important role? Are there social housing programs for lower income households?

Even if we are on the right direction to develop human-centred urban areas by facilitating caring activities, the chrono-urbanism needs to set the ground on aspects that specifically target caring activities, given the multiple impacts it has on everyday mobility, the unequal distribution of tasks among the population and its related socioeconomic consequences. Therefore, it is necessary to provide alternatives to measure progress on this important issue. City planning should respond to these issues to promote equity. On the other hand, an intersectional approach to understand how different personal identities interact in our daily day activities in the cities is necessary. 


# References

::: {#refs}
:::

<!-- `r word_comment("If you're interested in how the unweighted results look like for the same maps see the appendix. I've also calculated the cycling travel matrices as well as the 'no dependent' i.e., the same analysis but now only considering 3 care categories: health, errand, and grocery. For the time being though let's focus on this.", " ")` -->

# Appendix

```{r plot-figure10, out.width=600, out.height=600}
#| label: fig-Fig10
#| fig-cap: "A bivariate map of the cumulative opportuninities map (aggregated per category at DA) and the diversity measure. Amenity diverse = diviersity index above the region median. Caring/Complete = accessibility value above the region median. The Basemap shapefiles are retrieved from the 2021 Canadian census [@governmentofcanadaCensusPopulation2023], the Open Data Hamilton Portal [@opendatahamiltonCityBoundary2023] and the USGS [@greatlakesUSGS2010]."
knitr::include_graphics(glue::glue(here::here(), "/figures/all_cats_bivariate_entropy_copp_plot.png"))
```