---
title: "rejected_text"
format: html
---

## rejected from intro 

(e.g., Paris mayor Anne Hidalgo proposed plan in her 2020 re-election campaign [@teixeiraClassifying15minuteCities2024], neighbourhood guidance in Shanghai, China [@weng15minuteWalkableNeighborhoods2019])


The "15-Minute City" has been recently adopted by leaders as a way to promote human-scaled cities. As defined in @moreno_introducing_2021, the 15-Minute City is a urban planning model based in chrono-urbanism, a theory that emphasizes the positive impact on quality of urban life when urban space becomes multi-rythmic [related to chronotopy @mulicekUrbanRhythmsChronotopic2015]. Multi-rhythmic urban space is in opposition to industrial Fordism and single-use zoning [@mulicekUrbanRhythmsChronotopic2015; @morenoVilleQuartHeure2016]; the segregation of urban time and space for individual uses and mobility. The 15-Minute City thus one that accommodates the ability to reach essential destinations such as shops, work, and healthcare within a walkable radius. It would allow individuals to reclaim time spent on car mobility and give way to sustainable modes; prompting urban spaces that are responsive to human needs and environmental sensibilities [@Allam2022]. The 15-Minute City generated much interest in planning practice and public discourse, especially after Paris mayor Anne Hidalgo proposed a 15-minute city plan in her 2020 re-election campaign. Though the term 15-Minute City is relatively new, neighbourhood planning is not (e.g., Perry's Neighbourhood Unit Concept (NUC) from the 1930s [@kissfazekasCircleParadigms15minute2022], 15-minute neighbourhoods in China (X)). A critique of neighbourhood planning approaches have been their top-down nature [@talenSocialSciencePlanned2017]. The 15-Minute City esposes to be different, however top-down planning questions of how to equitable select travel-time thresholds (15 minute, or something else?), by what sustainable mode, to what destinations, and for whom continue to plague proximity-based urban planners.


While much care labour takes place in the private and feminised sphere, completing care activities requires one to be mobile. In other words: caring for oneself and others requires trips to, amongst other places, grocery stores, day cares, elder care, and health care destinations. 




@moreno_introducing_2021 considers four dimensions on how the 15-minute neighbourhood framework: density (in terms of people per km$^2$); diversity (including mixed-land use and diversity of people); the temporal and spatial proximity to essential services (residential areas, work, commercial areas, education centres, health facilities and other basic institutions) and digitalisation. 



<!-- These critiques were supported by case studies of planned neighbourhoods as... lead to social segregation of classes, destination functions, and automobile use. From it's onset, it was an efficient way to develop housing after the Second World War. The emphasis of proximity to some 'daily essentials' but disconnection from specialised activites (e.g., work) paved the way for car-based subrbia.  by capital of cities. Can the neighbourhood form might be seen as a vehicle for capitalist exploitation, gentrification and displacement??? The place of neighbourhood planning is actively being debated: can it's social crtiques be addressed in the new form as a "15-Minute City"? This study seeks to contribute to this decades old debate. -->

As related to Mobility of Care, care trips and hence care destinations are of key focus. While using entropy to measure the diversity of care destinations within 15-minute neighborhoods has yet to be explored, it theoretically complements the 15-Minute City concept.

The entropy measure has been widely used to characterise the diversity of land-use mix in early work by [@frankLinkingObjectivelyMeasured2005] based on the species diversity Shannon-Wiener index that expresses relative evenness throughout a sequence (see review in @whittaker1972evolution). It has also been used in social sciences to comprehend the occurrence of social events in a structured manner specially with regard to life trajectories [@ritschardSequenceAnalysisRelated2018]

A variety of tools have been applied in the literature to examine the 15-Minute City. However, the question of _what destinations matter_  still demands further examination. The following sub-section bridges the concept of the 15-Minute City with another normative framework: Mobility of Care.

Before discussing 'what outcome', we first ask 'what amenity' as the focus of this study. Often, accessibility analysis studies have focused on a single amenity type (X), often employment destinations and often focused on a scale beyond the neighbourhood. However, due to 15-Minute City's emphasis on proximate-level service provision, recent academic work on the topic has employed accessibility analysis to account for multiple 'local' amenities types, but seldom included employment. We argue that the inclusion of employment opportunities into the 15-Minute City conceptualisation is critical to it's definition. Why? We echo critics made by planners and social scientistics in decades past: neighbourhoods that did not mix work with service and culture within the neighbourhood, paved the way for car-based suburbia: cars allowed for the connection of neighbourhoods with their 'lower-order' amenities to areas with 'high-order' destinations like work (X). The lower- and higher- order framing of these destinations supported the continued segregation of urban life across many dimensions (X).

Combing all amenity types into the 15-Minute City concept is also a matter of gender equality. There is a branch of research that encompases this critique: Mobility of Care. This framework encourages the explicit consideration of destination types that serve unpaid labour (care destinations) into transport planning (as opposed to only considering destinations with paid work). Not considering these care destinations have travel time (and hence quality of life) implications on those who do care activites, commonly women and especially disadvantaged women. The consideration of _what_ amenity type is a gender issue. 

From this position, this study aims to holistically investigate the ease of interacting with proximate amenities (accessibility) from the perspective of the 15-Minute City and the Mobility of Care framework. As mentioned, accessibility analysis have typically focused on employment (X) and recent 15-Minute City flavoured analysis focus on 'lower-order' amentities. We focus on a case study of the mid-sized City of Hamilton, Canada, and aim to classify areas by what degree they may be considered a '15-Minute Neighbourhood" based on the relative quantity of amenity access and diversity of access. We offer a novel methodology of holistically considering all proximate-amenities, offer empirical findings, and contribute theoretically to the ongoing debate on the definition of a 15-Minute City by foregrounding care destinations. 


The objectives of the paper are as follows:

1)	To examine the potential accessibility to, and diversity of, care destinations in the case study. We identify to what degree are neighbourhoods caring and/or complete. To do so, we make use of self organizing maps (SOM), an unsupervised machine learning data-dimensionality reduction technique to identify caring and complete neighbourhoods.
2)	We then apply the same SOM methodology to determine profiles of population groups to understand who enjoys these neighbourhoods and who does not.





In this way, destinations outside of the home involved in caring activities has important implications for spatial and transportation planning; for all people but especially for all women and those from lower income households who complete the most mobility of care trips [@ravensbergen2023exploratory]. 


 In fact, early feminists in the United States had already observed that domestic labor was the single most relevant cause of women’s inequality in society [@gordon_womans_1976; @haydenGrandDomesticRevolution1982]. .... Sánchez de Madariaga not only shows how important these mobility of care trips are, but also highlights the ways in which “mobility of care is systematically under-represented in any analysis of urban transport” (p. 37). Transport surveys and tools do not directly capture mobility of care, which re-enforces the idea that these trips are not a significant part of daily mobility. .... In this way, destinations outside of the home involved in caring activities has important implications for spatial and transportation planning; for all people but especially for all women and those from lower income households who complete the most mobility of care trips [@ravensbergen2023exploratory]. 


Care also has important implications for spatial and transportation planning. Indeed, researchers from various fields have highlighted that western modern industrial societies have imposed a physical separation of the private sphere from the public one, splitting production and reproduction, public and private spaces or simply work and home, concentrating on the economic outcome of being mobile [@cresswell_gendered_2008; @lawWomenTransportNew1999; @levy2013travel; @little_gender_1994; @tronto_toward_1990]. Men tend to be associated with legal rights, paid labour, and the public sphere while woman are associated with the private sphere, dependents, family duties, personal relations and unpaid labour [@tronto_toward_1990]. In this line, @cresswell_gendered_2008 described how masculinity has been constructed as mobile and active, whereas femininity is static. As an example of this dichotomy, transport planning has historically framed non-work/school trips as 'non-mandatory' activities, downgrading these activities in comparison to mandatory trips like work and school (XXX). In more recent years, planners have begun to resist the spatial separation of the public and private sphere to promote for sustainable cities. Proximity planning has become more normalized, a popular example is the 15-minute neighbourhood i.e., neighbourhoods where residents can reach all necessary destinations within a 15-minute active transport trip. The 15-minute neighbourhood concept, however, has yet to foreground care.




<!-- The spatial structure of cities and regions has been described in different ways and its advantages and disadvantages has been discussed with regard to a wide variety of issues, for example, regarding health, the environment, economic, social and equity impacts. After the traditional monocentric vision of the city with a core Centre Business District (CBD), other alternatives appeared, especially due to the expansion of cities caused by the automobile and motorized transport. The polycentric urban planning emerged based on the idea of sub-centers (or multiple business centers) with their own services and functions, well-connected among other surrounding areas, where the links between land-use and transport were optimised. Economic activities (such as retail or commercial office development) and job centers played an important role in shaping the spatial structure [@schneider_transit_1981; @sweet_are_2017]. This concept may have different interpretations depending on the spatial and geographical scales [@hall_polycentricity_2009]. -->


Within the backdrop of urban planning and ‘mobility of care’, we connect our work to the 15-minute neighbourhood concept. 

Both the 15-minute neighborhood and accessibility measures are flexible enough to consider all destination types holistically, however, gendered examinations have been lacking in the literature. For instance, accessibility analyses have historically been employment-centric even when adopting a gendered perspective [@paez_jobs_2013]. Some 15-minute city analyses focus on care destinations, but they seem to not put care at the forefront, or to consider but one care destination (e.g., [@hosford15minuteCityReach2022] on grocery stores). Though few 15-minute neighbourhood analyses are explicitly feminist, their potential to build feminist communities have been noted outside of academia. For instance, Women in Urbanism have noted how women may benefit more than men from this 15-minute approach to planning. For instance, women tend to walk and use transit more than men, as well as to make more frequent, but shorter trips -and often as part of a chain (i.e., with stops along the way). These trip characteristics are all well served in 15-minute frameworks. 15-minute neighbourhoods also promote walkability, which results in more “eyes on the streets” and its associated safety, an important enabler of mobility for women and girls. Finally, because women tend to earn less money than men and cars are a large financial burden, planning paradigms that reduce car dependency, such as 15 minute cities, may on average benefit women more than men [@Bertolo2023].

Furthermore, mobility of care trip characteristics differ from other trips, notably those of the commute to work. For instance, many of these trips are made through trip-chains, i.e., when a non-work-related stop is made during the commute. Women have been found to do so more than men, which is understood as being due to their need to balance unpaid-household serving care and paid employment (e.g., picking up groceries or children on the way home from work) [@ravensbergen2023exploratory; @scheinerWomenComplexDaily2015]. 

-- 
# possible discussion

Second.5. the discusison of self-sorting: single parents provide more care and hence may be more likely to choose, if they are able, more accessible neighbourhoods than the rest of the population in the same household income grouping. In a Canadian and U.S. context, households with children may locate to catchments with 'better' schools, often more suburban with lower potential access scores overall. Furthermore, statistics show that in the last years, the number of single parents families in Canada has increased. In 2022, there were about 1.84 million single parents compared to about 1.56 million in 2010 and 1,3 million in 2001 (https://www.statista.com/statistics/443342/single-parent-families-in-canada/ & https://www.justice.gc.ca/eng/rp-pr/jr/jr12/p5a.html#ftn4). Furthermore, Ontario is the Canadian province where most of the lone parent families live (749,769 families, https://www.statista.com/statistics/446667/lone-parent-families-in-canada-by-province/). Most lone-families are headed by female-lone parent. (AS TO ADD IN HAMILTON SPECIFIC NUMBERS)

In car dependent cities, cycling commute rates tend to be low. Emerging evidence has found that they are even lower for care trips than for work trips. For instance, in Montréal it is estimated that 2% of work trips completed by adults are done by bicycle, while only 1.1% of care trips are done using this sustainable travel mode [@ravensbergen2023exploratory]. Qualitative research has also found that in car-dependent contexts, many consider it inappropriate and unsafe to conduct escort trips by bicycle [@ravensbergenVelomobilitiesCareLowcycling2020; @sersliRidingAloneTogether2020]. Then again, other studies have uncovered the joys inherent in travelling by bicycle with children [@RN14]. This qualitative evidence has also found that the bicycle is excellent for trip-chaining owning to its flexibility and ease of parking, and that grocery shopping can be done by bicycle, though this can require some additional materials (e.g., paniers) and competencies (e.g., knowing where to put your groceries)[@ravensbergenVelomobilitiesCareLowcycling2020]. Research in the Netherlands, a high-cycling context, has found that parents can and do use their bicycles to complete mobilities of care. Indeed, one study compared women with children’s bicycle behaviours to those without children and found that the having a child did not change the amount of cycling done, but the characteristics of cycling trips: mothers’ trips tended to center around their childrens’ needs [@RN15]. Taken together, it is evident that the bicycle has potential to serve mobility of care: it is flexible, low-cost, and used for these trips in many contexts. Other barriers to this travel mode seems to exist in low-cycling contexts. 

The difference in uptake of public transport for work compared to care trips is staggering. In Montréal, it is estimated that 18.7% of trips to work are made by transit. In contrast only 5.8% of care trips are made using transit [@ravensbergen2023exploratory]. While research examining accessibility to care destinations by transit exists (e.g., [@RN16;@RN17;@RN18;@RN19]), little research examines why few people use this sustainable transport mode for mobility of care. Though transit use for mobility of care is low, one study found that is it is significantly higher for low-income women. Indeed, approximately 16% of low-income women in Montreal were found to use transit for care trips (compared to 4% of high income women) [@ravensbergen2023exploratory]. In many cities, public transport is not designed to meet household demands in a convenient way [@RN20;@grantsmithManagingChallengesCombining2016]. Given how low transit use is for care trips amongst the general population, it is possible that many of these low-women using transit are transit-dependent. In other words, they may be using transit for care trips because they do not have access to a car, and not because transit is their preferred travel mode. 

It appears that the car is the preferred travel mode for mobility of care. In one study, car use was higher for mobility of care trips than the commute to work (82.3% vs. 74.7%) [@ravensbergen2023exploratory]. Further, the rate of car use is higher amongst trips that are part of a chain than those that are not (85.28% vs. 78.09%) [@ravensbergen2023exploratory]. Some studies found that a car is viewed as the easier way to incorporate unpaid mobility of care into one’s daily travel as it provides convenience and mitigates concerns about traffic and stranger danger [@maciejewskaHaveChildrenThus2019, @carverParentalChauffeursWhat2013]. Others, however, have found that women are more likely to complete household-serving travel duties by alternative modes than men due to their lower car access [@RN24;@RN25]. It seems that car use may be preferred for mobility of care trips, but planning mobility of care for the car may leave those who do not have access to a car, many of them low-income women, with difficult mobility of care trips. Studies also observe disparities regarding gender mobility and type of activities, depending on the area of residence; women living in the city center, where more transport systems and activities are available, use more sustainable travel modes than the ones living in the outskirts, which tend to rely more on cars to carry out daily activities [@marquet2014walking; @mejia2021mobility; @montero2023role; @montero2023applying]. Outer areas are not as dense as in the downtown, and are basically car-oriented developments. Thus, studies observe that in the outer areas gender equity is confronted to sustainable mobility. 

The discussion around mobility of care and travel mode may also be relevant to larger spatial trends, including gendered spatial entrapment. Feminist geographers have noted for decades the spatial impact of women’s disproportionate household-serving labour. A relationship has been noted whereby many women’s household responsibilities results in job searches that emphasize job opportunities that are near the home and part-time or with flexible hours to allow for care responsibilities. This leads to a lower probability of securing higher paid, and often further afield, employment [@hansonGenderMobilityNew2010;@lawWomenTransportNew1999;@RN28;@RN29]. This trend has been found to be surprisingly resilient, with improvements only taking places slowly [@RN30]. Evidence has also shown that this spatial entrapment can be more in car dependent contexts, as access to cars impacts employment rates [@RN31], and is lower amongst women, and especially women of colour [@RN32].

Taken together, there appears to be a preference for travel by car for mobility of care trips in many geographical contexts. This is problematic, however, in at least two major ways. First, cars are less sustainable than other modes of transport. Normalizing car use for care trips does not advance the low-carbon transport transition that most cities are aiming for (at least in principle!). Since care trips tend to be short, they are also perfect candidate to be completed by bus, bicycle, or by foot. That is, if these alternative and sustainable modes are planned with care in mind. Second, there is equity concerns with assuming people are accessing care destinations by car. Many people, and low-income women in particular, do not have access to a car. Some are also experiencing a degree of spatial entrapment. Having safe and convenient access to care destinations using alternatives to the private car is especially important for those who must rely on these options regardless of whether they are well planned. 

Even if we are on the right direction to develop human-centred urban areas by facilitating caring activities, the chrono-urbanism needs to set the ground on aspects that specifically target caring activities, given the multiple impacts it has on everyday mobility, the unequal distribution of tasks among the population and its related socioeconomic consequences.

The process of gentrification and displacement are trends also occurring in Hamilton. On the horizon of rising rents, there is a risk of displacement among more vulnerable households, who may be replaced by individuals with higher incomes who can afford different transport modes, with more flexible schedules and are more likely to use ICT arrangements (including telework, e-shopping, and other online services). While, lower income households may be particularly vulnerable to locating in areas with worse accessibility to care alternatives, which could have negative consequences for these households and for the society as a whole.

Even if we are on the right direction to develop human-centred urban areas by facilitating caring activities, the chrono-urbanism needs to set the ground on aspects that specifically target caring activities, given the multiple impacts it has on everyday mobility, the unequal distribution of tasks among the population and its related socioeconomic consequences. Therefore, it is necessary to provide alternatives to measure progress on this important issue. City planning should respond to these issues to promote equity. On the other hand, an intersectional approach to understand how different personal identities interact in our daily day activities in the cities is necessary.



Even if we are on the right direction to develop human-centred urban areas by facilitating caring activities, the chrono-urbanism needs to set the ground on aspects that specifically target caring activities, given the multiple impacts it has on everyday mobility, the unequal distribution of tasks among the population and its related socioeconomic consequences.
Therefore, it is necessary to provide alternatives to measure progress on this important issue.
This work leaves many doors open for further research. For example,
•	the role that gender plays in these households.
•	Job arrangements per type of household. Some studies observe that working from home arrangements are more likely available for higher education attainment households (likely to have higher income), and families with children, but also living in neighbourhoods with good accessibility,
•	Finally, an intersectional approach to understand how our personal identities interact in our access to essential activities and caring services.
City planning should respond to these issues to promote equity. On the other hand, an intersectional approach to understand how different personal identities interact in our daily day activities in the cities is necessary.

As in other case studies, essential opportunities and services tend to concentrate on Hamilton central. Therefore, it is more likely to carry out daily activities, including caring ones, in this area using sustainable modes such as walking or biking.

However, this study highlights that there seems to be a spatial mismatch between dwelling availability, family profile and eventually (housing and mobility) affordability. At the neighbourhood level, the concentration of opportunities for all category types is in Hamilton Central, but households with children may not live there.

Richer families may concentrate in either neighborhoods classified with poor accessibility: either “Somewhat caring and not complete” or “Not caring and not complete. However, these families are likely to have different mobility alternatives available. The literature shows that higher income households have a lower mobility share compared to lower income households. Furthermore, they are more likely to be assisted by different service arrangements (i.e. e-shopping, service deliveries, taxis and MaaS services, working from home arrangements,,…).

Our results show a different story for low-income households.  If they can afford it, these households will prefer to live in very caring and very complete neighbourhoods, which will ease their accessibility to daily activities and essential services and therefore, avoiding social exclusion. In this respect, one-parent families are more likely to be time poor. For them, living in low access/incomplete suburbs increases their chancesof  falling into social exclusion. Furthermore, school closures or any other type of relocation of essential services puts more pressure on their shoulders.

## about employment.

Though not considered, employment is a promiment destination type not considered within this work. However, as mentioned, care activites are often trip-chained. therefore, the availability of care destinations around work also matter. This is for future investigation.

In Canada, women have become more educated and achieved increased participation in the labour market [@government_of_canada_gender_2023]. In fact, the educational level of women has surpassed most men [@government_of_canada_women_2023]. Reports observe that the gender pay gap has narrowed but disparities persist [@drolet_intersectional_2023;@pay_equity_office_gender_2023]. For instance, in the province of Ontario women still make 87 cents on average for every dollar made by a man and by all reports women of inter-sectional marginalized identies are more impacted [@pay_equity_office_gender_2023]. Similarly to other countries worldwide, the occupational segregation remains and women are segregated into lower paid job classes and there is a dominance of women in minimum wage jobs [@drolet_intersectional_2023]. Furthermore, women are underemployed, and make up most part-time work. In this respect, the lack of childcare alternatives enforces this obligatory part-time work [@equal_pay_coalition_myths_2022]. All researchers and organisations agree that an intersectional approach is of utmost importance to better plan policies to enforce gender equality. 


# possibel conclusions


Cities have simplified individuals as a universal and neutral entity. As a result, public spaces are differently experienced depending on our multiple identities. Indeed, asymmetric in the distribution of time of citizens is deeply anchored with the public and private spaces organization within cities. In this light, we bring together the literatures on 15-minute cities and mobility of care by focusing on the potential access to destinations of care through an empirical case study of the mid-sized city of Hamilton, Canada. We acknowledge that caring is composed by actors and activities but aim to spotlight the role that the urban environment plays on our daily lives, on easing our participation and on exercising our rights to the city as individuals with multiple identities. Our work hopes to contribute to a fruitful discussion that may trigger urban and transport policy interventions that increase gender-equity in caring activities among all actors.






```{r plot-figure6-create, eval=FALSE}
plot1 <- ggdraw() + draw_image(glue::glue(here::here(), "/figures/access_sc_plot_2.png"))
plot2 <- ggdraw() + draw_image(glue::glue(here::here(), "/figures/access_sc_plot_1.png"))
plot3 <- plot_grid(plot1, plot2) 

ggsave(glue::glue(here::here(), "/figures/access_sc_plot_combo.png"), dpi=300)
```

```{r plot-figure6, out.width=800, out.height=800}
#| label: fig-Fig6
#| fig-cap: "The 5 resulting superclusters from the SOM output (left) and the associated dendrogram structure (right)."
#| 
knitr::include_graphics(glue::glue(here::here(), "/figures/access_sc_plot_combo.png")) 
```

<!-- Describing @fig-Fig9, the most valuable variable for splitting the data was the proportion of single family households and % of LICO prevalence in the DA. The the terminal nodes represented as bar columns depict the percentage composition by supercluster. For the bar on the right, the majority of the parcels are F superclusters. Following the tree diagram, these F superclusters are described by having less than 15.8% single parent families (this is below the median of `r df_DA$One_parent_families |> median(na.rm=T) |> scales::percent()`). The bar on the left is mostly composed of parcels that are labeled A+. These A+ parcels are located in DAs greater than 15.8% single parent families and % LICO prevalence of 6.5% and greater (this is above the median of `r df_DA$LICO |> median(na.rm=T) |> round() |> paste0("%")`). In sum: residences in parcels with exceptionally complete and caring access tend to be lower-income and single-parent families, while residences in parcels with not caring are not complete access tend to have a lower presence of single parent households. Households with a higher proportion of single-family households (15.8% and greater) where LICO is lower have a mixed label membership (bar column in @fig-Fig9).  -->


<!-- As a tool to visualise the variables that describe the superclusters, a Decision Tree is fit with the DEPENDENT TOTALS, ERRAND TOTALS, GROCERY TOTALS, HEALTH TOTALS, and CATEGORY DIVERSITY scores. Aligning with the descriptive summaries of the superclusters provided, describing the tree-diagram from left to right (roughly from the most caring & complete, to the least), the first three terminal nodes (represented by bar columns that visualise the proportion of supercluster membership at that node) are mostly A+ and A superclusters, and are defined by their high dependent-centric access (about median and above), and high diversity score. A+ is differentiated from A by its exceptionally high health-centric totals while A most just has excellent scores. Columns four, six and seven are dominated by the presence of B+ and B parcels. On average the B+ parcels provide access to an average level of dependent-centric destinations, so the B+ parcels with greater than 6.5 dependent-centric destinations are found in the more left-most columns, and those below this split on the right-most. The far right-most column is dominated by the presence of F parcels, defined by lower DEPENDENT TOTALS and LOW DIVERSITY. It is interesting to note the 5th column, which is defined by D parcels but also the most amount of A- parcels than any other column. A- and D parcels are both high in dependent-centric destinations and low in diversity (though D is much lower than A-). These two parcels again, as shown in @fig-Fig7, can be thought of as parcels in neighbourhoods containing a high amount of some caring destinations, but lacking others, indicating high efficiency for improvement potential.  -->

<!-- ```{r plot-figure7, out.width=600, out.height=300} -->
<!-- #| label: fig-Fig7 -->
<!-- #| fig-cap: "The Decision Tree demonstrating variables that define the superclusters." -->
<!-- knitr::include_graphics(glue::glue(here::here(), "/figures/access_sc_profiles_usingSCVARS.png")) -->
<!-- ``` -->


<!-- D (Nothing much: Not caring & not complete) (C7) -->

<!-- B- (Below average, but dep- heavy: Somewhat caring & not complete) (C4) -->
<!-- B (Average: Caring & complete) (C5) -->
<!-- B+ (Above average: Caring & very complete) (C6) -->

<!-- A- (Excellent, but dep- heavy: Very caring & not complete (C3) -->
<!-- A (Excellent: Very caring & very complete) (C2) -->
<!-- A+ (Exceptional: very caring & very complete) (C1) -->

OLD RANDOM HIGHLIGHTS:

- The 15-Minut City have re-emerged as a popular concept in urban-planning city-planning; emphasizing the access to urban amenities by foot or bicycle. However, historically, neighbourhood planning processes have been top-down; FMC is linked to the Neighbourhood Unit Concept of the 1920s, largely critized for supporting segregated neighbourhoods, leaving a void in neighbourhood planning concepts that were filled by auto-centric suburban sprawl. Regardless, fundamental concepts within FMC remain popular: people identify with neighbourhoods, see cities through that lense. These ideas have in part re-emerged in the popularity of 15-minute neighbourhoods. 

- But how do we uplift the positive, popular, aspects of FMN without opening it up to the critics of the NUC? In this work, we turn to feminist geographies and the Mobility of Care conceptualisation and apply it to FMC. We argue, conventional planning practice is masculinist. This can threaten the FMC concept by shaping how 'important' amenities are conceptualised thus favouring work- and leisure- centered amenities in cities that are increasingly capital-driven. Though FMN is fundamentally based in bottom-up collaborative planning processes, we exist in the context of planning pracice which is largely masculinist. 

- FMN are for people - people do unpaid labour, those people are often women and intersectionally women-of-colour and low-income. However, this work should be for all people. Cities should be designed with this conceptualisation in mind. This is to combate the growing capitalisation of cities. 

- Our paper offers this nuanced understanding of FNC -- calling for them to center Care Destinations as important amenity types with the explicit conceptualisation that their destination type matters especially for women (low-income and of colour) who interact with them most, but also for all people. Their explicit positioning also can indirectly combat finacialisation of cities and co-opting of the FMC concept to one that centers capital-generating destination types like work and prohibitive leisure. Not to say those destination types should or cannot co-exist with Care destinations -- they should -- but markets are already incentivizing them. The market does not speak for non-economically geberating destination types, so we must explicitly call for them and support them within the FNC framework. 
- 
- 